Unveiling Robustness Threats by Mining the Exception Stack Traces of over 600 Android Projects

### 2014-09-06 android-stacks.tex
Shortened abstract.
And fixed a couple of English issues.

\begin{abstract}
The number of mobile apps is increasing in a daily rate. If on the one hand they
extend phones capabilities far beyond of the basic calls, on the other hand they have to 
cope with an increasing number of exceptional conditions 
(e.g., faults in underlying middleware or hardware; memory and battery restrictions). 
Therefore, mechanisms for exception detection and handling are not an optional add-on but a 
fundamental part of such apps. However, studies have shown that such mechanisms are the less 
understood and tested parts of the system. As a consequence they may affect the system the 
other way around, making the system even less robust, leading to failures such as uncaught exceptions 
and unintended handling (i.e., exceptions being mistakenly caught by existing handlers).
This paper reports a study which mined the stack traces embedded on 
159,048 issues reported on Android projects hosted in GitHub (482 projs.) 
and Google Code (157 projs.). 
The goal of this study was to: investigate the whether the exception stack traces
 can reveal common characteristics on the exception-related code that
may be favoring the introduction of the aforementioned failures, which would be responsible for 
decreasing the application robustness. We call such characteristics ``fragilities" of the exception-related code.
Overall 6,005 exception stack traces were extracted, and analyzed in 
combination with source code and bytecode analysis. 
Some fragilities consistently detected in this study include: 
(1) undocumented runtime exceptions thrown by third party code;
(2) undocumented checked exceptions thrown by a JNI interface function;
(3) unexpected exception wrappings (e.g., Errors being wrapped in checked exceptions),
revealing that Java hybrid exception model is not fully used according to its purpose.
Moreover, about 50\% of all analyzed projects reported at least one NullPointerException
as the root cause of an exception stack trace.
This study provides insights on such fragilities and the robustness threats  
they impose not only to Android apps but also to other systems 
based on Java exception model.

@@

To guide this investigation we compiled general guidelines on how to use
exceptions proposed by Gosling~\cite{gosling2000java},
Wirfs-Brock~\cite{wirfs2006toward} and Bloch~\cite{bloch2008effective}.
Then, using a custom tool called ExceptionMiner,
 which we develop specifically for this study, we mine stack traces from the issues reported 
on 482 Android projects hosted in GitHub and 157 projects hosted in Google Code.
Overall 159,048 issues were analyzed and 6,005 stack traces were extracted from them.
The exception stack trace analysis was augmented by means of byte code 
and source code analysis
 with additional exception-related information
extracted from the JDK, Android Platform and the application layer.
%Manual inspection was used to leverage the understanding 
%of exception stack traces and support further discussions and insights.

@@

@@

\noindent \fbox{ \begin{minipage}{0.96\columnwidth} \emph{
 267 exception stack traces were caused by ``undocumented runtime exceptions" thrown by libraries or the Android Platform.
Methods defined on 11 android packages signaling direct instances of RuntimeException.
}
\end{minipage}}

>>>>

\begin{abstract}
%The number of mobile apps is increasing in a daily rate. If on the one hand they
%extend phones capabilities far beyond of the basic calls, on the other hand they have to 
%cope with an increasing number of exceptional conditions 
%(e.g., faults in underlying middleware or hardware; memory and battery restrictions). 
%Therefore, mechanisms for exception detection and handling are not an optional add-on but a 
%fundamental part of such apps. However, studies have shown that such mechanisms are the less 
%understood and tested parts of the system. As a consequence they may affect the system the 
%other way around, making the system even less robust, leading to failures such as uncaught exceptions 
%and unintended handling (i.e., exceptions being mistakenly caught by existing handlers).

This paper reports a study which mined the Java stack traces included in
159,048 issues reported on Android projects hosted in GitHub (482 projects) 
and Google Code (157 projsects). 
The goal of this study is to investigate  whether the exception handling logic on these stack traces
 can reveal common characteristics on the exception-related code that
are particularly fault prone and may lead to a decrease in application robustness.
Overall 6,005 exception stack traces were extracted, and analyzed in 
combination with source code and bytecode analysis. 
Key problems identified in this study include: 
(1) undocumented runtime exceptions thrown by third party code;
(2) undocumented checked exceptions thrown by a JNI interface function;
(3) unexpected exception wrappings (such as Errors being wrapped in checked exceptions),
indicating that Java's hybrid exception model is not fully used in accordance with its purpose.
Moreover, about 50\% of all analyzed projects reported at least one NullPointerException
as the root cause of an exception stack trace, suggesting the presence of programming mistakes.
This study provides insights on such fragilities and the robustness threats  
they impose not only to Android apps but also to other systems 
based on Java exception model.

@@

To guide this investigation we compiled general guidelines on how to use
exceptions proposed by Gosling~\cite{gosling2000java},
Wirfs-Brock~\cite{wirfs2006toward} and Bloch~\cite{bloch2008effective}.
Then, using a custom tool called ExceptionMiner,
 which we develop specifically for this study, we mine stack traces from the issues reported 
on 482 Android projects hosted in GitHub and 157 projects hosted in Google Code.
Overall 159,048 issues were analyzed and 6,005 stack traces were extracted from them.
The exception stack trace analysis was augmented by means of byte code 
and source code analysis
 with additional exception-related information
extracted from the Android Platform and Android applications.

@@

%Manual inspection was used to leverage the understanding 
%of exception stack traces and support further discussions and insights.

@@

\noindent \fbox{ \begin{minipage}{0.96\columnwidth} \emph{
The study revealed: 267 exception stack traces were caused by ``undocumented runtime exceptions" thrown by libraries or the Android Platform; 
several methods defined on 11 android packages signaling direct instances of RuntimeException.
}
\end{minipage}}
